Noisy Work: A Review of The Uncertainty Visualisation Literature

### 2024-01-30 paper.qmd
added a philosophy part to the uncertainty definition part, and did a lot of the general attitudes towards uncertainty

---
title: "The Noisy Work of Uncertainty Visualisation"
author: Harriet Mason
bibliography: references.bib
date: last-modified
format: pdf
editor_options: 
  chunk_output_type: console
---

@@

Examples of Taxonomies used to untanlge the features of uncertainty:
- aleatory uncertainty (unavoidable randomness), epistemic uncertatiny (uncertainty about structure, still quantified), ontological uncertainty (uncertain about entire modelling process). @Spiegelhalter2017
- measurement (uncertainty on a single varible), spatial (uncertainty in spatial data), temporal (uncertainty in a time data)
- aleatory (inherrent randomness in a forecast) vs epistemic (uncertainty in structure and parameters of statistical models) vs ontological (uncertainty about the entire modelling process as a description of reality)
- scientific methods (variance, distribution) vs human judgments (disagreement among parties)
- Random vs systematic uncertainty
- statistical (can provide a continuous pdf) vs bounded (know values will fall in a certain range) uncertainty
- accuracy vs precision
- Uncertainty from discovery, assessment and execution
- risk vs uncertainty
- "known knowns, known unknowns, and unknown unknowns", (risk, uncertainty, ignorance)
- Source of uncertainty (model structure, parameters, forcing functions, intial state, model operation) and diagnostic (past/current) vs prognostic (forecasting)
- uncertainty is organised according to Satience (legal moral, societal, institutional, proprietary, situational) and level of severity (high, medium, low)
- Level of precision in epistemic uncertainty: 1) numbers given to appropriate levels of precision 2) a distribution range 3) a measure of statistical significance 4) verbal quantifiers to numbers 5) refusal to give a number unless the evidence is good enough @Spiegelhalter2017

@@

[Illustration of the connection between the data generating process and the location of the uncertainty analysis. Several examples are collected from utypo and Munzner2009. ](location.png){#fig-location}

@@

Mapping uncertainty into the 3D space defined by location, level, and nature can expand beyond the narrow case of statistical modelling it was defined for. *@___* included "consensus" as a taxonomy element, however it was not included in the final definition, but it can easily be added as a location that would align with the communication stage of the data pipeline. Several authors [*@_+_@Spiegelhalter2017*] considered the method by which uncertainty was quantified (e.g. by a single value, a function, simulated outcomes, etc), which can be be thought of as a more precise categorisation of the level axis. This flexibility means that the use of this framework is not limited by a specific case and should be used going forward to define what aspect of the uncertainty space a contribution exists in.

*Should I add in a bit where uncertainty papers are organised according to which uncertainties they visualise or should I leave that until the lit review section?*

@@

# 3.General Attitude Towards Uncertainty + Its communication
I dont think this is worth a section but it is mentioned a lot in .
- the role of unertainty visualisation is to communicate accurate information
- problem of communicating to a group thing (avoid communicating inaccurate information)
- how people feel about uncertainty is irrelevant
  - natural disasters make people uncomfortable but to include it in a conversation about whether or not we should express it is rediculous.
- Numerical risks: Use absolute risks (but relative risk when catastrophic event), for single unique event use percent chance, express change as a proportion, frequency, or percentage (and be clear about reference class), avoid framing bias by showing risk with and without the outcome, keep the denominator fixed when making comparisons, explicit about time interval, For more knowledgeable audiences consider providing epistemic uncertainty and confidence in analysis @Spiegelhalter2017

@@

- compares a bunch of stuff to the chance of getting struck by lightning since they are on a common scale (which is kind of funny) @Spiegelhalter2017
- hard to communicate small risks because people struggle to discrimintate between 1 in 10 000 and 1 in 100 000. Common solution is to use a logarithmic scale (if drawn vertically these are known as risk ladders). design of scale important because risk is associated with position rather than magnitude @Spiegelhalter2017
- heath had a wide range of different metrics they adopted to communicate chronic risk of adverse events in the future @Spiegelhalter2017
- using words to communicate numerical risk is generally discouraged and using words without numerical interpretation can lead to lower accuracy and higher risk aversion @Spiegelhalter2017
- the Affect heuristic comes to play in how people understand risk @Spiegelhalter2017
- avoids the topic of risk at times of accute crisis because numerical risk would be inappropriate @Spiegelhalter2017
- positive vs negative framing matters, can try to avoid it @Spiegelhalter2017
- no preference between percent or frequency although (small interaction with numeracy skills though) @Spiegelhalter2017
- Need careful specification of the reference class (e.g. % chance of rain should be fractions of situations like this in which is rains - over both time and location) @Spiegelhalter2017
- Need to make time interval clear (50% each year) @Spiegelhalter2017
- understanding better in a table than text @Spiegelhalter2017
- if expressing as a frequency, note that larger numerators suggest a larger risk, exteme version of this is ignoring the denominator all together @Spiegelhalter2017

@@

# 4 LITERATURE SUMMARY
- methods used in uncertainty visualisation evaluation remain ad hoc. studies approach issue from a usability perspective instead of asking WHY representations DO or DO NOT work. Studies do not follow any methodology commonly agreed upon. @Kinkeldey2014
- Lack of formalisation and eigour in empirical methdos is an issue that is much broader than uncertainty visualisation (extends to info-vis and related domains). Therefore this paper is a starting point, not a definitive summary @Kinkeldey2014
- The question in these papers is often "does method a work better than method B" which is an engineering approach, where the toal is to improve a particular product rather than create general principals @Kinkeldey2014
- Authors then try to generalise beyond the specific constraints of the test even though the framing does not allow for it @Kinkeldey2014
- Alternatively, we should ask how/why does method A work better than method B, which grounds work in perceptual and cognative theory and provides a framwork @Kinkeldey2014

@@

# 4.1 Plot Comparison Papers
# 4.1.1 The problem of incomparable plots
Visual inference can be seen as a process that combines multiple fields, from mathematics to psychology, to convey meaning. Unfortunately which tasks belong to mathematics, computer science, or psychology is poorly defined. The uncertainty visualisation taxonomies make the blur between these fields. A visualisation is, in a lot of ways, a statistic, or at the very least statistics are calculated prior to the visualisation being calculated, however this may be invisible to the user. A density plot does not just "appear" from the data, usually a smoothing function (that should be calibrated according to the sample size however how this should be done is a different matter of dispute) generates an estimated density function (statistics) which is then depicted on a plot using a line (design or computer science) which is then converted into information in your head using perceptual tasks and heuristics (psychology).

@@

- Suggests that it is straightforward to show a value but much more complex to show uncertainty @Hullman2016 (I think this is interesting in relation to how people SEE uncertainty visualisation. Why is uncertainty hard to visualise?)
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016
- When asking questions, researchers should focus on events that can be repeated (because of a misunderstanding of confidence intervals) @Hullman2016
- Uncertainty defined as a by product of the process of making a visualisation, from data collection to commucating @Hullman2016
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016
- communication tasks: This category comprises map reading tasks involving data and uncertainty value retrieval (which location is most uncertain?). For tasks from this category, visualisation techniques can be chosen following the traditional rules from cartography. @Kinkeldey2014 (basically saying why dont you just use what is already here if you are treating uncertainty as a signal, which I agree with) 
- papers usually have same variance if there is a signal question, and use the mean as a proxy to answer the question. Really should ask a changing signal question where the signal IS impacted by the uncertainty information.
- Weirdly When setting their ground truth it seems like a lot of papers EXPECT participants to completely ignore uncertainty information as noise. Like the participants are actually considered to be incorrect if they dampen their signal understanding because of noise considerations which I find FUNNY. 
- I think the questions where uncertainty is a signal, not a noise, allow the authors of visualisation experiments to ignore the context and motivation that usually governs a visualisation.
# 4.?? Tasks and Motivating Questions (uncertainty about.. .what?)
- "cannot assess the quality of risk communication unless the objectives are clear" circles back to motivation. He assumes we are fulfilling a duty to inform. @Spiegelhalter2017
## 4.?? What are you uncertain about? A mathematical considerations
- Should add plot comparison mathematical framework example here
- take a plot and work through the entire example as a flow chart thing
## 4.1?? In other statistics
What is meant by "uncertainty" may seem obvious to some, but when you attempt to quantify or visualise it you will quickly find yourself asking, "uncertainty about... what?". Do you mean uncertainty on an estimate? On a forecast? How many steps ahead is this forecast? Are we only considering the uncertainty in the estimate or in the parameters or are we considering the possibility of measurement error or biased inputs? Signal and noise can only be untangled in the presence of a motivating question.

@@

## 4.1?? General task notes
- Mentioned alleatory vs epistemic uncertainty, in the visualisation literature alleatory uncertainty is more commonly the focus. also mentions dependence and joint pdfs as a consideration @Hullman2016
- Most studies involve intrinsic approaches such as colour, transparency, and those that used extrinsic techniques typically used glyphs, error bars, grid based techniques, or contouring @Kinkeldey2014
- most studies used coincident approaches, static visualisations, are applied to a specific domain and dont have generalised results @Kinkeldey2014
- The tasks for participants were most commonly value retrival (some where uncertainty and data value are retrived separately, others where they are some separately). also also be asked to extract a relative value @Kinkeldey2014
- How does the complexity of uncertainty relate to the cateogies of user and task @Kinkeldey2014
- Several studies probide evidence that the usability of uncertainty representations can be highly user and task dependent @Kinkeldey2014
- The nature of the task plays an important role for the usability of uncertainty techniques, this may explain many of the insonsitent outcomes from the studies under review since two studies assessing sifferent techniques and user groups are not comparable @Kinkeldey2014
- Doing a follow up publication to discuss issues with reasoning and decision making based on uncertainty visualisations which is not considered or discussed in this paper @Kinkeldey2014
- Main take away is that "we need to systematize future empirical studies on uncertainty visualisation to better enable comparison and generalisation of the findings"
- One way to advance this goal is a taxonomy, however existing taxonomies are focused on data types, uncertainty categories, and representation types @Kinkeldey2014
- he lists a bunch of tasks that were tested with uncertainty visualisationsat the end @Kinkeldey2014

>>>>

---
title: "Noisy Work: A Review of The Uncertainty Visualisation Literature"
author: Harriet Mason
bibliography: references.bib
date: last-modified
format: pdf
editor_options: 
  chunk_output_type: console
---

@@

Examples of Taxonomies used to untanlge the features of uncertainty:

- aleatory uncertainty (unavoidable randomness), epistemic uncertatiny (uncertainty about structure, still quantified), ontological uncertainty (uncertain about entire modelling process). @Spiegelhalter2017
- measurement (uncertainty on a single varible), spatial (uncertainty in spatial data), temporal (uncertainty in a time data)
- aleatory (inherrent randomness in a forecast) vs epistemic (uncertainty in structure and parameters of statistical models) vs ontological (uncertainty about the entire modelling process as a description of reality)
- scientific methods (variance, distribution) vs human judgments (disagreement among parties)
- Random vs systematic uncertainty
- statistical (can provide a continuous pdf) vs bounded (know values will fall in a certain range) uncertainty
- accuracy vs precision
- Uncertainty from discovery, assessment and execution
- risk vs uncertainty
- "known knowns, known unknowns, and unknown unknowns", (risk, uncertainty, ignorance)
- Source of uncertainty (model structure, parameters, forcing functions, intial state, model operation) and diagnostic (past/current) vs prognostic (forecasting)
- uncertainty is organised according to Satience (legal moral, societal, institutional, proprietary, situational) and level of severity (high, medium, low)
- Level of precision in epistemic uncertainty: 1) numbers given to appropriate levels of precision 2) a distribution range 3) a measure of statistical significance 4) verbal quantifiers to numbers 5) refusal to give a number unless the evidence is good enough @Spiegelhalter2017

@@

![Illustration of the connection between the data generating process and the location of the uncertainty analysis. Several examples are collected from utypo and Munzner2009. ](location.png){#fig-location}

@@

Mapping uncertainty into the 3D space defined by location, level, and nature can expand beyond the narrow case of statistical modelling it was defined for. *@___* included "consensus" as a taxonomy element, however it was not included in the final definition, but it can easily be added as a location that would align with the communication stage of the data pipeline. Several authors [*@_+_@Spiegelhalter2017*] considered the method by which uncertainty was quantified (e.g. by a single value, a function, simulated outcomes, etc), which can be be thought of as a more precise categorisation of the level axis. This taxonomy also ignores contextual information about the data, such as spatial or temporal considerations. This flexibility means that the use of this framework is not limited by a specific case and should be used going forward to define what aspect of the uncertainty space a contribution exists in.

The taxonomies that cannot be contained within the @utypo taxonomy are not a new uncertainty dimensions, but rather they are conflating adjacent information with uncertainty. A good example of this is the distinction between spatial, temporal, and estimated uncertainty. When we create statistical models and quantify uncertainty, the spatial or temporal aspects are often boilled away [@Otsuka2023], and the only element of it that remains is the inexchangeability of our data. For this reason, there realistically is no such thing as "spatial" uncertainty in the same sense that there is no such thing as "photograph" uncertainty or "coin toss" uncertainty or "baby birth" uncertainty. Maps and photographs boil down to the same statistical objects in the same way that coin tosses and baby births boil down to the same statistical objects. Once we have removed the physical nature these objects are near identical when considered statistically. Of course the nature of them can change based on what we choose to observe (i.e. if we consider a coin landing on its side to be an outcome or if we consider intersex babies as an outcome in births) but this means the contextual information of our data, e.g. whether or not it is spatial or temporal, is only of importance at two stages of the data analysis pipeline. It is first important when we *boil it down* to our data through *observation* and again when we *map our statistical information back to its real world information* to *communicate* it. The only way the context uniquely contributes to uncertainty is in the "boil down" stage, however spatial or temporal uncertainty never refers to that specific cause of uncertainty. It is typically in reference to measurement error, data aggregation, statistical modelling, or other steps that occur *after* the boil down stage, and are therefore not unique to spatial or temporal data. While the spatial or temporal features *may* become relevant at the final communication stage, however this has little do with quantifying uncertainty, and the difficulties in communicating a complicated statistical model through a complicated map is already well discussed in the visualisation literature, and in later sections we will discuss how this concept is related to uncertainty visualisation.

In this sense we are not proposing that the current conceptualisation of data according to its physical properties is necessarily wrong, or that it leads to incorrect conclusions, we are merely suggesting for a different way of slicing the metaphysical cake of uncertainty that better aligns with what is *actually happening* when we perform a statistical analysis. This metaphysical cake is illustrated below as @fig-cake. In changing  what the actual boundaries are when we work with these problems, we improve our understanding of what we are actually doing when we conceptualise uncertainty, and in the process allow for more creative solutions creativity (e.g. by borrowing concepts from spatial visualisation to visualise uncertainty in photographs generated). However, since uncertainty is almost synonymous with the concept of "statistics", this adjustment in how we conceptualise and organise statistical methods would likely benefit other domains within the field.

![The metaphysical cake we are slicing differently](metaphysicalcake.jpg){#fig-cake}

It is important to note that this taxonomy is the closest we can get to mathematical definition of uncertainty. While individual *expressions of uncertainty*, such as a confidence interval or a PDF are mathematically defined, the term *uncertainty* cannot be defined mathematically because it is a layperson word that has been adopted by statistics to describe any deviation from complete determinism, whether mathematically defined or not. While the broader field of uncertainty visualisation does need to be more closely inspected, specifically through the scope of mathematical definitions, this issue cannot be solved with a more specific definition of uncertainty through mathematics. Instead or establishing a new definition of uncertainty that is strict enough to prevent the current array of misunderstandings, we should focus on inspecting how the current literature *deviates* from established mathematical definitions, why these deviations occur, and understand how these deviations can lead to larger problems in uncertainty communication.

@@

# 3.General Attitude Towards Uncertainty (Uncertainty Communication Boradly)
It would be impossible to discuss uncertainty visualisation without mentioning the concept of uncertainty communication and discussing how visualisation fits into that broader scope. Uncertainty visualisation is related to uncertainty communication in the same way that visualisation more broadly is related communication more broadly. Visualisation is fast, it can communicate more complicated ideas that what can be done with text alone, and

*psychology of uncertainty communication*

It is difficult to read a paper that discusses uncertainty that does not make a nod towards uncertainty bias. In the case of uncertainty visualisation, these nods are not always relevant. Similarly to how contextual information is conflated with uncertainty when we try to define it, human attitudes to risk are often commented on in the discussions of uncertainty visualisation. Ironically, despite being mentioned in a large proportion of papers that discuss uncertainty risk aversion, has very to do with the psychology of uncertainty. As a matter of fact, it has so little to do with uncertainty communication that we would not even mention it in this paper were it not for the fact that ignoring it would suggest we had missed an important section of the literature.

Risk-aversion is the tendancy of people to prefer outcomes of low uncertainty to those outcomes with high uncertainty despute the outcomes with high uncertainty having a higher or equal expected outcome.

Uncertainty information is typically included in some form of statistical communication to offer some transparency to the viewer. Data analysis is a long and complicated process, so much so that even experts in the field can find themselves lost with a single incorrect assumption, so uncertainty can help pull back the curtain on this process and prevent people from understanding it as a human agnostic, entirely deterministic process. When science authors discuss "risk avoidance" in the uncertainty communication conversations, it holds with it some implication that heuristics such as risk avoidance are relevant in every conversation about uncertainty communication, however that is not true.

Risk and uncertainty are not perfect synonyms of one another. @Spiegelhalter2017 makes the distinction clear by specifying *risk* as *a precise random event* (such as a coin flip) while uncertainty is imprecision (such as the estimated probability associated with a coin that might be biased). Additionally, risk avoidance is observed when people have a choice between a *deterministic* outcome and a *risky* event that has higher expected value. How this psychological heuristic is therefore related to all uncertainty visualisation despite uncertainty communication being different in both the scenario (i.e. it is highly unlikely that we need to communicate a choice between a perfectly deterministic and an uncertain event) and in the type of uncertainty (since risk and uncertainty are not synonyms).

The regular conflation of risk and uncertainty despite clear and specific differences, without any explanation of how these terms are related, implies that a large portion of uncertainty literature is simply reporting information from any research that is uncertainty adjacent without considerations for the specifics of how these concepts would be related to uncertainty.

It is not even clear that this behaviour is entirely irrational either, considering risk aversion is, at its core, a trade off between bias and variance. If risk-aversion is truly "irrational" behaviour, every statistical textbook that discusses the use of a biased but consistent estimator in a fite sample case (a common example would be specific cases of the maximum likelihood estimator) should immediately toss those chapters out and the authors should feel embarassed for that *clearly* irrational line of thinking. Of course I am making a joke but this does highlight a clear problem in the way risk-aversion is discussed in the literature. Discussing risk-aversion as though it is some diversion from a perfectly rational, mathematical choice, implies there is *no value to a decrease in variance* which is not even true *within mathematics*. The concept of risk aversion treats risk as "noise", as though it is something that only exists to be ignored, but if the common place use of biased MLE's is an indication of anything, it is that *ignoring uncertainty* is the irrational choice.

Selecting a biased is technically statisticians doing "risk aversion", but when used to ironically when it is used to select models it is not viewed as an irrational human fault.

This issue, where psychological considerations seem to be entirely devorced from what is *actually* being discussed extends to the conversation around trust as well. The idea of measuring or being concerned about audiences perception of trust, without first establishing if what we are communicating is *trustworthy*, . The goal of uncertainty visualisation should not be to be trustworthy, but rather to be transparent (which can facilitate trust).

I dont think this is worth a section but it is mentioned a lot in .
- problem of communicating to a group thing (avoid communicating inaccurate information)
- how people feel about uncertainty is irrelevant
  - natural disasters make people uncomfortable but to include it in a conversation about whether or not we should express it is rediculous.
- Numerical risks: Use absolute risks (but relative risk when catastrophic event), for single unique event use percent chance, express change as a proportion, frequency, or percentage (and be clear about reference class), avoid framing bias by showing risk with and without the outcome, keep the denominator fixed when making comparisons, explicit about time interval, For more knowledgeable audiences consider providing epistemic uncertainty and confidence in analysis @Spiegelhalter2017

@@

- compares a bunch of stuff to the chance of getting struck by lightning since they are on a common scale (which is kind of funny) @Spiegelhalter2017
- hard to communicate small risks because people struggle to discrimintate between 1 in 10 000 and 1 in 100 000. Common solution is to use a logarithmic scale (if drawn vertically these are known as risk ladders). design of scale important because risk is associated with position rather than magnitude @Spiegelhalter2017
- using words to communicate numerical risk is generally discouraged and using words without numerical interpretation can lead to lower accuracy and higher risk aversion @Spiegelhalter2017
- the Affect heuristic comes to play in how people understand risk @Spiegelhalter2017
- avoids the topic of risk at times of accute crisis because numerical risk would be inappropriate @Spiegelhalter2017
- positive vs negative framing matters, can try to avoid it @Spiegelhalter2017
- no preference between percent or frequency although (small interaction with numeracy skills though) @Spiegelhalter2017
- Need careful specification of the reference class (e.g. % chance of rain should be fractions of situations like this in which is rains - over both time and location) @Spiegelhalter2017
- Need to make time interval clear (50% each year) @Spiegelhalter2017
- understanding better in a table than text @Spiegelhalter2017
- if expressing as a frequency, note that larger numerators suggest a larger risk, exteme version of this is ignoring the denominator all together @Spiegelhalter2017

@@

# 4 LITERATURE SUMMARY
- methods used in uncertainty visualisation evaluation remain ad hoc. studies approach issue from a usability perspective instead of asking WHY representations DO or DO NOT work. Studies do not follow any methodology commonly agreed upon. @Kinkeldey2014
- Lack of formalisation and eigour in empirical methdos is an issue that is much broader than uncertainty visualisation (extends to info-vis and related domains). Therefore this paper is a starting point, not a definitive summary @Kinkeldey2014
- The question in these papers is often "does method a work better than method B" which is an engineering approach, where the toal is to improve a particular product rather than create general principals @Kinkeldey2014
- Authors then try to generalise beyond the specific constraints of the test even though the framing does not allow for it @Kinkeldey2014
- Alternatively, we should ask how/why does method A work better than method B, which grounds work in perceptual and cognative theory and provides a framwork @Kinkeldey2014

@@

# 4.1 Plot Comparison Papers
# 4.1.1 The problem of incomparable plots
Visual inference can be seen as a process that combines multiple fields, from mathematics to psychology, to convey meaning. Unfortunately which tasks belong to mathematics, computer science, or psychology is poorly defined. The uncertainty visualisation taxonomies make the blur between these fields. A visualisation is, in a lot of ways, a statistic, or at the very least statistics are calculated prior to the visualisation being calculated, however this may be invisible to the user. A density plot does not just "appear" from the data, usually a smoothing function (that should be calibrated according to the sample size however how this should be done is a different matter of dispute) generates an estimated density function (statistics) which is then depicted on a plot using a line (design or computer science) which is then converted into information in your head using perceptual tasks and heuristics (psychology).

@@

- Suggests that it is straightforward to show a value but much more complex to show uncertainty @Hullman2016 (I think this is interesting in relation to how people SEE uncertainty visualisation. Why is uncertainty hard to visualise?)
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016
- When asking questions, researchers should focus on events that can be repeated (because of a misunderstanding of confidence intervals) @Hullman2016
- Uncertainty defined as a by product of the process of making a visualisation, from data collection to commucating @Hullman2016
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016
- communication tasks: This category comprises map reading tasks involving data and uncertainty value retrieval (which location is most uncertain?). For tasks from this category, visualisation techniques can be chosen following the traditional rules from cartography. @Kinkeldey2014 (basically saying why dont you just use what is already here if you are treating uncertainty as a signal, which I agree with) 
- papers usually have same variance if there is a signal question, and use the mean as a proxy to answer the question. Really should ask a changing signal question where the signal IS impacted by the uncertainty information.
- Weirdly When setting their ground truth it seems like a lot of papers EXPECT participants to completely ignore uncertainty information as noise. Like the participants are actually considered to be incorrect if they dampen their signal understanding because of noise considerations which I find FUNNY. 
- I think the questions where uncertainty is a signal, not a noise, allow the authors of visualisation experiments to ignore the context and motivation that usually governs a visualisation.

# 4.?? Tasks and Motivating Questions (uncertainty about.. .what?)
- "cannot assess the quality of risk communication unless the objectives are clear" circles back to motivation. He assumes we are fulfilling a duty to inform. @Spiegelhalter2017
## 4.?? What are you uncertain about? A mathematical considerations
- Should add plot comparison mathematical framework example here
- take a plot and work through the entire example as a flow chart thing
## 4.1?? In other statistics
What is meant by "uncertainty" may seem obvious to some, but when you attempt to quantify or visualise it you will quickly find yourself asking, "uncertainty about... what?". Do you mean uncertainty on an estimate? On a forecast? How many steps ahead is this forecast? Are we only considering the uncertainty in the estimate or in the parameters or are we considering the possibility of measurement error or biased inputs? Signal and noise can only be untangled in the presence of a motivating question.

@@

## 4.1?? General task notes
- heath had a wide range of different metrics they adopted to communicate chronic risk of adverse events in the future @Spiegelhalter2017 (use this for motivation based statistics since they define new statistics based off what they need)
- Mentioned alleatory vs epistemic uncertainty, in the visualisation literature alleatory uncertainty is more commonly the focus. also mentions dependence and joint pdfs as a consideration @Hullman2016
- Most studies involve intrinsic approaches such as colour, transparency, and those that used extrinsic techniques typically used glyphs, error bars, grid based techniques, or contouring @Kinkeldey2014
- most studies used coincident approaches, static visualisations, are applied to a specific domain and dont have generalised results @Kinkeldey2014
- The tasks for participants were most commonly value retrival (some where uncertainty and data value are retrived separately, others where they are some separately). also also be asked to extract a relative value @Kinkeldey2014
- How does the complexity of uncertainty relate to the cateogies of user and task @Kinkeldey2014
- Several studies probide evidence that the usability of uncertainty representations can be highly user and task dependent @Kinkeldey2014
- The nature of the task plays an important role for the usability of uncertainty techniques, this may explain many of the insonsitent outcomes from the studies under review since two studies assessing sifferent techniques and user groups are not comparable @Kinkeldey2014
- Doing a follow up publication to discuss issues with reasoning and decision making based on uncertainty visualisations which is not considered or discussed in this paper @Kinkeldey2014
- Main take away is that "we need to systematize future empirical studies on uncertainty visualisation to better enable comparison and generalisation of the findings"
- One way to advance this goal is a taxonomy, however existing taxonomies are focused on data types, uncertainty categories, and representation types @Kinkeldey2014
- he lists a bunch of tasks that were tested with uncertainty visualisationsat the end @Kinkeldey2014
